---
title: "Untitled"
author: "Christopher Hayes, David Hakula, Filip Mellgren"
date: '2019-04-13'
output:
  html_document:
    code_folding: hide
    df_print: kable
    highlight: zenburn
    theme: readable
  pdf_document: default
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r cars, include = FALSE}
library(rio)
library(tidyverse)
library(AER)
library(ggthemes)
library(stargazer)
library(reshape2)
library(xtable)
#library(plm)
library(glue)
library(zoo)
```
# Step 0
```{r}
df <- import("Cars.dta")
df <- df %>% mutate(yearcountry = paste(year, country)) %>% as_tibble()
```

# 1 Summary statistics
Show and discuss summary statistics of the car market in the five EU countries. Focus on what you believe are key variables given the questions at hand.

Could be the case: "Note that the panel is unbalanced because most models are not available throughout the entire period or in all countries."
```{r}
# Create two functions that are used throughout the analysis to create 
# descriptive tables

# TODO: Add necessary summary stats. Do we need percentiles?
summary_stats <- function(x){
  x <- as.matrix(x)
  
  summary <- rbind(mean(x), sd(x, na.rm = TRUE), min(x, na.rm = TRUE), 
                   max(x, na.rm = TRUE))
    return(summary)
}

# Takes above and turns it into a nice table
summary_table<- function(summary){
  summary <- as_tibble(summary) %>% t() %>% round(2) %>% as_tibble(rownames = "Variable") %>% rename ("Mean" = V1, "S.e." = V2, "Min." = V3, "Max." = V4)
  return(summary)
}
```


```{r}
# Summary statistics to Q1.
# TODO: Group by country? If so, stack them, keep the same columns
descriptive_table <- df %>%
  select(country, year, domestic, qu, price, princ, horsepower, fuel, pop) %>% 
  mutate(pop = pop/1000000) %>%
  map(function(x) summary_stats(x)) %>% summary_table() %>% mutate(country = "All")

for (c in 1:5) {
 descriptive_table <-  rbind(descriptive_table, df %>% filter(country == c) %>% 
  select(country, year, domestic, qu, price, princ, horsepower, fuel, pop) %>% 
    mutate(pop = pop/1000000) %>%
  map(function(x) summary_stats(x)) %>% summary_table() %>% mutate(country = c))
}



tmp_table <- descriptive_table %>% arrange(Variable) %>% 
  mutate("se" = paste(S.e., ")", sep = ""),
    "Mean_se" = paste(Mean, se, sep = " (")) %>% 
  select(Variable, Mean_se, country) %>% # Add min and max later for simplicity
  spread(key = country, value = Mean_se)

tmp_table <- rbind(tmp_table, tmp_table) %>% arrange(Variable) %>% 
  filter(Variable != "country") %>% 
  select(Variable, All, "1", "2", "3", "4", "5") %>% rename("Belgium" = "1", 
                                                            "France" = "2",
                                                            "Germany" = "3",
                                                            "Italy" = "4",
                                                            "UK" = "5")

tmp_table_means <- tmp_table[seq(1, nrow(tmp_table), 2), ] %>% mutate(All = word(All,1),
                                                   Belgium = word(Belgium),
                                                   France = word(France,1),
                                                   Germany = word(Germany,1),
                                                   Italy = word(Italy,1),
                                                   UK = word(UK,1))

tmp_table_se <- tmp_table[seq(1, nrow(tmp_table), 2), ] %>% mutate(All = word(All,2),
                                                   Belgium = word(Belgium,2),
                                                   France = word(France,2),
                                                   Germany = word(Germany,2),
                                                   Italy = word(Italy,2),
                                                   UK = word(UK,2))
minmax_table <- df %>%
  select(country, year, domestic, qu, price, princ, horsepower, fuel, pop) %>% 
  mutate(pop = pop/1000000) %>%
  map(function(x) summary_stats(x)) %>% summary_table() %>% 
  select(Variable, Min., Max.) %>% filter(Variable != "country")

                                                   
final_table <- rbind(tmp_table_means, tmp_table_se) %>% full_join(minmax_table) %>% arrange(Variable) %>% 
  select(Variable, All, Min., Max., Belgium, France, Germany, Italy, UK)

final_table
```

```{r}
# Correlation matrix to Q1.
# Strictly necessary?
cormatdf <- df %>% select(price, qu, year, domestic, horsepower, fuel, princ, weight)
cormat <- round(cor(cormatdf),2)

melted_cormat <- melt(cormat)

melted_cormat %>% ggplot(aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + theme_economist() + scale_colour_economist() +
  labs(title = "TILE", y = " ", x = " ",
       caption = "caption") +
  scale_fill_gradient2(low = "#336666", high = "#664033", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  geom_text(aes(Var1, Var2, label = value), color = "black", size = 4)
```
Excluded co (model code), segment (ordinal variable), firm, country, because they are nonsensical
Also height and width were excluded, because they were not deemed as important
Pop was excluded because it is not a feature related to the cars

Discuss the difference in prices across countries. Look at UK mean prices for example.

Tendency to buy domestic cars

Mention the data set that we have access to (briefly)

# 2 Hedonic price regression

```{r}
# TODO: add the correct variables to X
# TODO: Squeeze the latex table
# TODO: Add country f.e. to the normal 
# TODO: Add segment dummies?

X <- glue("year", "horsepower", "fuel", "width","height", "domestic", "year", .sep = "+")

formula <- glue("log(price) ~", X)
lm.fit <- lm(formula, data = df)

```
```{r}
X_fe <- glue(X, "country1", "country2", "country3", "country4", .sep = "+")
fe_formula <- glue("log(price) ~", X_fe, "factor(co)", .sep = "+")
fe.fit <- lm(fe_formula, data = df)

stargazer(lm.fit, fe.fit, type = "text", omit = "factor", omit.yes.no =
c("Yes", "No"))
```


# 3 Market shares

Calculate the following market shares: (a) the market share of each
car model, and (b) the market share of each car model in its segment.
Discuss your findings. Assume that the total market size is the number
of households, approximated by total population divided by 4

Comment: "In practice, the potential market size in a given year may be lower because cars are durable and consumers who just purchased a car may not consider buying a new one
immediately."

```{r}
# TODO: maybe take the five highest and five lowest per a and b. Most popular cars and least popular car

#a
df <- df %>% mutate(msize = pop/4, 
              mshare = qu/msize,
              ln_mshare = log(mshare))
# b
df <- df %>% group_by(segment, yearcountry) %>% mutate(tot_s = sum(qu)) %>%
  ungroup()
df <- df %>% mutate(mshare_s = qu / tot_s,
                    ln_mshare_s = log(mshare_s))

# Table 1
df %>% mutate(mshare = mshare *100,
              mshare = mshare_s *100) %>% 
  select(mshare, mshare_s, ln_mshare, ln_mshare_s) %>% 
  map(function(x) summary_stats(x)) %>% summary_table()

# market shares by firm over year, car models and countries
# Likely not necessary
df %>% group_by(firm) %>% summarise(mean = mean(mshare)*100, se = sd(mshare),
                                    min = min(mshare), max = max(mshare))

```



# 4 Logit demand
Supposed to pool the data?

Estimate demand using a logit specification (Berry, 1994). Use different estimators such as OLS and fixed effects. Use your estimated
coefficients to calculate own-price elasticities. Interpret your results.
Are there any interesting differences across firms?

The fixed effects may be on a model level!!!!
```{r}
df <- df %>% group_by(yearcountry) %>% mutate(tot_qu = sum(qu),
                                        oshare = 1 - tot_qu/msize) %>% ungroup()

df <- df %>% mutate(ln_dmshare = log(mshare) - log(oshare))
```

```{r}
# Adds price to the X
X <- glue("price", "year", "horsepower", "fuel", "width","height", "domestic", "year", .sep = "+")
X_fe <- glue(X, "country1", "country2", "country3", "country4", .sep = "+")

formula <- glue("ln_dmshare ~", X)
fe_formula <- glue("ln_dmshare ~", X_fe, "factor(co)", .sep = "+")
lm.logit <- lm(formula, data = df)
fe.logit <- lm(fe_formula, data = df)

# TODO: do we present this or is it enough with the elasticities?
stargazer(lm.logit, fe.logit, type = "text", omit = "factor")
```

```{r}
# TODO: what coefficient do we use to estimate the elasticities? Maybe both
alpha_1 <- lm.logit[["coefficients"]][["price"]]
df <- df %>% mutate(own_price_elasticity = alpha_1 * (1 - mshare)*price)

summary_stats(df$own_price_elasticity) %>% summary_table()
```

```{r}
# What's this good for?
df <- df %>% group_by(firm) %>% mutate(avg_e_jj_l = mean(own_price_elasticity)) %>% ungroup()

df <- fastDummies::dummy_cols(df, select_columns = "firm")


```


# 5 Nested logit demand
Estimate demand using a nested logit specification where segment is
defined as the nest (Berry, 1994). Use different estimators such as OLS
and fixed effects. Use your estimated coefficients to calculate own-price
elasticities. Discuss your results.

```{r}
df <- df %>% mutate(ln_mshare_s = log(mshare_s))


# Adds ln_mshare_s to the X
X <- glue("price", "year", "horsepower", "fuel", "width","height", "domestic", 
          "year", "ln_mshare_s", .sep = "+")
X_fe <- glue(X, "country1", "country2", "country3", "country4", .sep = "+")

formula <- glue("ln_dmshare ~", X)
lm.nest <- lm(formula, data = df)
fe_formula <- glue("ln_dmshare ~", X_fe)
fe.nest <- lm(fe_formula, data = df)
```

```{r}
# TODO: again, do what elasticfity do we calculate?
alpha_2 <- lm.nest[["coefficients"]][["price"]]
sigma <- lm.nest[["coefficients"]][["ln_mshare_s"]]

df <- df %>% mutate(e_jj_nl = alpha_2 * (1/(1-sigma) - (sigma/(1-sigma)) * mshare_s  - mshare)*price)

# TODO: fix output
df %>% select(e_jj_nl,own_price_elasticity) %>%
  map(function(x) summary_stats(x)) %>% summary_table()
```

```{r}
# Calculates average average elasticity within nest for each firm
df <- df %>% group_by(firm) %>% mutate(avg_e_jj_nl = mean(e_jj_nl)) %>% ungroup()
```


```{r}

# TODO: Make it pretty!
df %>% group_by(firm) %>% summarise(mean = abs(mean(own_price_elasticity))) %>% 
  ggplot(aes(x = reorder(firm, -mean), y = mean)) +
  geom_bar(stat="identity") +
  theme_economist() + scale_fill_economist() + scale_color_economist() +
  labs(title = "Elasticities")

df %>% group_by(firm) %>% summarise(mean = abs(mean(e_jj_nl))) %>% 
  ggplot(aes(x = reorder(firm, -mean), y = mean)) +
  geom_bar(stat="identity")
```


# 6 Instruments

Goldberg, Verboven 2001 section 4.2 provides some insight: variables that shift the producers' supply relations, but are excluded from the demand equations, make natural instruments
 
 
Suggest two types of instruments that could be used to solve the endogeneity problem of the price variable. Why do you propose these instruments?

Want to find something that shifts the supply curve:
Things that change the marginal cost:

* price on car paint
* Strikes specific to the car industry or where strike wage remains. 

To solve the endogeneity problem of the price variable, we need instruments that shifts the supply in market, without affecting the demand.

As a first instrument, we suggest price on car paint. This clearly affects the price through the supply side, by increasing the marginal costs of the supplier. The price on car paint hence should be relevant (note that this is testable, given data on price of car paint). It is unlikely that the price on car paint affects the market price through any channel on the demand (or any other channel for that matter) and hence it is likely that the instrument would also be valid. 

Furthermore, we suggest to use data on strikes specific to the car industry. This should increase the production costs and hence shift supply, making the instrument relevant (not again that this is testable, given access to data). The validity of the instrument relies on the assumption that the strike does not affect the strikers demand for cars in such a way that total market demand is affected. We deem the validity as likely: the total population share working in the car industry is usually rather low on a national level. Also, strike funds are often utilized during strikes to compensate strikers for the absent wages. 


# 7 Nash Bertrand
Assume Nash-Bertrand competition in prices on the supply side. Calculate the implied marginal costs, markups and Lerner index based on
the estimates from the nested logit specification in [5]. Present firmlevel averages of price, marginal cost, markup and the Lerner index.

Do you think your results are economically meaningful? Explain and discuss.

```{r}
# TODO: Note that markups are negative
# markups
df <- df %>% mutate(pderiv = alpha_2 * mshare*((1/(1-sigma))-(sigma/(1-sigma))*mshare_s-mshare),
                    markup = -mshare/pderiv, # "-" to get the right sign
                    mc = price - markup,
                    lerner = markup/price)
# TODO: what is markup2? 

# summarise

```
```{r}
# firm level averages
df <- df %>% group_by(firm) %>% mutate(avg_markup = mean(markup),
                                 avg_mc = mean(mc),
                                 avg_lerner = mean(lerner),
                                 avg_price = mean(price)) %>% ungroup()

# Table for question 7. Pre merger descriptives. 
pricing_table <- df %>% select(firm, avg_price, avg_mc, avg_markup, avg_lerner) %>% 
  distinct(firm, .keep_all = TRUE) %>% arrange(-avg_price)

```


# 8 Merger
Note that Matilda made a mmistake in the lecture, she divided by 4 to get a 25% decrease.

```{r merger_function}
post_merger_df <- function(dataframe, ratio_mc){
  # this takes a data frame, creates new variables and returns the same data frame
  dataframe <- dataframe %>% 
    mutate(firmnew = if_else(firm == 15 & country == 3 & year >= 1998, 26, firm), 
           segment_new = if_else(co == 177 & country == 3 & year >= 1998, 4, segment), 
           qu_new = case_when( 
             co == 164 & country == 3 & year >= 1998 ~ qu*0.75, 
             co == 166 & country == 3 & year >= 1998 ~ qu*0.5, 
             co == 168 & country == 3 & year >= 1998 ~ 0, 
             TRUE ~ qu), 
           mshare1 = qu_new/msize)
  
dataframe <- dataframe %>% group_by(segment_new, yearcountry) %>% 
  mutate(tot_s1 = sum(qu_new), 
         mshare_s1 = qu_new / tot_s1) %>% ungroup()

# TODO: double check we should be using alpha_2 and not something else
dataframe <- dataframe %>% mutate(mc = mc * ratio_mc,
                                  pderiv_new = alpha_2 * mshare1*((1/(1-sigma))-(sigma/(1-sigma))*mshare_s1-mshare1),
                                  markup_new = -(mshare1 / pderiv_new),
                                  price_new = mc + markup_new,
                                  lerner_new = markup_new / price_new,
                                  dprice = (price_new-price)/price)

# By firm: avg_price, avg_mc, avg_markup, avg_lerner
dataframe <- dataframe %>% group_by(firm) %>% mutate(avg_markup_new = mean(markup_new),
                                 avg_mc_new = mean(mc),
                                 avg_lerner_new = mean(lerner_new),
                                 avg_price_new = mean(price_new)) %>% ungroup()


  return(dataframe)
}
```

```{r}
merger_df <- post_merger_df(df, 1)
```


```{r}
merger_df %>% filter(year == 1998, country == 3) %>%
  select(mshare_s, mshare_s1) %>% map(function(x) summary_stats(x)) %>% summary_table()
```
## Post merger
```{r}
# Compare per-merger and post-merger prices: 
# combine with table from 7
pricing_table_new <- merger_df %>% select(firm, avg_price_new) %>% 
  distinct(firm, .keep_all = TRUE)

full_join(pricing_table, pricing_table_new) %>% 
  mutate("Relative price change" = avg_price_new/avg_price-1) %>%
  select(avg_price, avg_price_new, "Relative price change", everything()) 
```



```{r}
ger_table <- merger_df %>% filter(year == 1998, country == 3) %>% drop_na() %>% 
  select(price, price_new, dprice) %>% map(function(x) summary_stats(x)) %>% summary_table() %>% cbind(name = "ger")

# GM
gm_table <- merger_df %>% filter(year == 1998, country == 3, firm == 15) %>% drop_na() %>% 
  select(price, price_new, dprice) %>% map(function(x) summary_stats(x)) %>% summary_table() %>% cbind(name = "gm")

# VW
vw_table <- merger_df %>% filter(year == 1998, country == 3, firm == 26) %>% drop_na() %>% 
  select(price, price_new, dprice) %>% map(function(x) summary_stats(x)) %>% summary_table() %>% cbind(name ="vw")

price_change_table <- rbind(ger_table, gm_table, vw_table) %>% select(Variable, Mean, name) %>% spread(key = Variable, value = Mean) %>% select(-dprice) 

price_change_table
```


Note that dprice does not correspond to (price_new - price) / price.

# 9 Efficient merger
Redo 8 with a changed mc. Tabulate all 5 segments.

```{r}
merger_df_9 <- post_merger_df(df, 0.8)
```

```{r}
merger_df_9 %>% filter(year == 1998, country == 3) %>%
  select(mshare_s, mshare_s1) %>% map(function(x) summary_stats(x)) %>% summary_table()
```

```{r}
# TODO: Add the five segments like the others were added
ger_table <- merger_df_9 %>% filter(year == 1998, country == 3) %>% drop_na() %>% 
  select(price, price_new, dprice) %>% map(function(x) summary_stats(x)) %>% summary_table() %>% cbind(name = "ger")

# GM
gm_table <- merger_df_9 %>% filter(year == 1998, country == 3, firm == 15) %>% drop_na() %>% 
  select(price, price_new, dprice) %>% map(function(x) summary_stats(x)) %>% summary_table() %>% cbind(name = "gm")

# VW
vw_table <- merger_df_9 %>% filter(year == 1998, country == 3, firm == 26) %>% drop_na() %>% 
  select(price, price_new, dprice) %>% map(function(x) summary_stats(x)) %>% summary_table() %>% cbind(name ="vw")

price_change_table_2 <- rbind(ger_table, gm_table, vw_table) %>% select(Variable, Mean, name) %>% spread(key = Variable, value = Mean) %>% select(-dprice)

price_change_table_2 <- price_change_table_2 %>% rename(efficient_price = price_new)
```

```{r}
# Merge the two price change tables
pre_post_table <- full_join(price_change_table, price_change_table_2)

pre_post_table <- pre_post_table %>% gather(price_type, price, -name) %>% 
  mutate(treatment = if_else(price_type == "price", 0, 1))
```


# 10 2 more features
We know what happens to production, but happens to sales?
What branding effects are there that might affect sales?

More nests?

The analysis conducted above focuses on the company level and production aspect of the merger. However, other important aspects of the businesses involved are likely to be affected as well. 

One aspect is sales and marketing. The car market is likely to be characterized by consumer loyalties, that perhaps is not captured in the product characteristics analysis made, above that could be affected by a merger: It is not curtain (as the above conducted analysis assumes) that the demand for a particular brand stays the same. E.g., some customers might value the “Americaness” and the American values that GM symbolizes. Such customers might be less willing to buy their models after a merger, where the GM production has been taken over by German Volkswagen. To the contrary, it could also create interesting co-branding opportunities. Such aspects of sales and marketing should hence also be considered in merger analysis. This could for example be done by incorporating these aspects into the model above, by scoring of these aspects of the products and estimation of how they would change with the merger through consumer surveys. 

Another aspect to be considered is the number of nests. We believe that a model using a richer variety of nests would more accurately capture the market behavior. [Filip: I don’t remember what we said here?]

Taxes varying across nation? or is it captured by country dummies?


```{r}
pre_post_table <- pre_post_table %>% select(-treatment) %>%  
  spread(price_type, price) 
```
```{r}
# TODO: make a similar one for segments
pre_post_table %>%
  ggplot(aes(x = as.factor("Pre merger"), y = price)) +
  geom_segment(aes(xend = as.factor("Post merger"), yend = price_new,color = "#336666")) +
  geom_segment(aes(xend = as.factor("Post merger"), yend = efficient_price, color = "#9ae5de")) +
  geom_point(size = 3) +
  geom_point(aes(x = as.factor("Post merger"),y = price_new), size = 3) +
  geom_point(aes(x = as.factor("Post merger"), y = efficient_price), size = 3) +
  theme_economist() + scale_fill_economist() + scale_color_economist() +
  annotate("text", x = 0.85, y = pre_post_table$price, label = as.character(pre_post_table$name)) +
  labs(title = "What are the effects of merger on prices?", y = "Average price",
       tag = "Graph X") +
  scale_x_discrete(name ="", limits=c("Pre merger","Post merger")) +
  scale_colour_manual(guide = "legend", name = "Merger type:",
         values =c("#336666"="#336666","#9ae5de"="#9ae5de"), 
         labels = c("Not cost reducing","Cost reducing")) +
  theme(plot.tag.position = "bottom") +
  scale_y_continuous(limits=c(10, 20))

  
```
```{r}
ggsave("merger_causal_effect.png", height = 20, width = 12, units = "cm")
```



