---
title: "Empirical workshop 4"
author: "Filip Mellgren, David Hakula, Chris Hayes"
date: "7 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, include = FALSE}
library(tidyverse)
library(rio)
library(readxl)
library(magrittr)
library(glue)
library(ggplot2)
library(AER)
library(ggthemes)
library(stargazer)
library(reshape2)
library(xtable)
#library(plm)
library(zoo)

```

```{r}
df <- import("jec.dta")
br1 = 28         # w28 1880
br2 = 11 + 3*52  # w11 1883
br3 = 26 + 3*52  # w26 1883
br4 = 12 + 6*52  # w12 1886
df <- df %>% rename(week = WEEK,
                    month = MONTH,
                    lakes = LAKES,
                    tqg = TQG,
                    po = PO,
                    gr = GR,
                    pn = PN)
df <- df %>% mutate(coll   = (po+pn)/2,
                    yrexact=(week/52)+1880,
                    year = floor(yrexact),
                    wy = ifelse(week %% 52 == 0, 52, week %% 52),
                    era   = case_when(
                      week < br1                ~ 1/5,
                      between(week, br1,   br2-1) ~ 2/5,
                      between(week, br2, br3-1) ~ 3/5,
                      between(week, br3, br4-1) ~ 4/5,
                      between(week, br4, 328) ~ 5/5),
                    gtr   = between(week, br1, br2),
                    nyc   = between(week, br2+1, br3),
                    cain  = between(week, br3+1, br4),
                    caout = between(week, br4+1,   328))
#count(df,pn-po)
```

```{r}
#ggplot(data=df, aes(x=yrexact, y=gr)) + geom_line(stat="identity")
# Scatter price vs qty (with different series for collusive periods, and perhaps different episodes (partial out lakes?))
library(reshape2)
ggplot(data=(df %>% select(yrexact,coll,gr,era) %>% melt(id="yrexact")), 
       aes(x=yrexact, y=value, colour = variable)) + geom_point()
# To find out info about theta, specific the supply equation. Summarise MC by weights of market share, 
ggplot(data=df, aes(x=tqg, y=gr+rnorm(1, mean = 0, sd = 0.01), colour=coll)) + geom_point() 
```

```{r}
qd    <- lm(log(tqg) ~ log(gr) + lakes + factor(month), data = df) # qty = f(price, lakes)
ps.po <- lm(log(gr) ~ log(tqg) + po + factor(era) + factor(month), data = df) # price = f(qty, lakes, era)
ps.pn <- lm(log(gr) ~ log(tqg) + pn + factor(era) + factor(month), data = df) # price = f(qty, lakes, era)
qd.coeff    <- summary(qd)[["coefficients"]]
ps.po.coeff <- summary(ps.po)[["coefficients"]]
ps.pn.coeff <- summary(ps.pn)[["coefficients"]]
```


## Overcoming endogeneity - IV specifications of supply and demand
Demand equation: qty = f(price, lakes, month). To overcome endogeneity of price, IV candidates are supply shifter: collusion and firm entry. 
```{r}
iv.qd.po <- ivreg(log(tqg) ~ log(gr) + lakes + factor(month)
               | factor(era) + lakes + factor(month), data = df) # collusion and entry IV for price
iv.qd.pn <- ivreg(log(tqg) ~ log(gr) + lakes + factor(month)
               |  factor(era) + lakes + factor(month), data = df) # collusion and entry IV for price
#iv_qd_sum <- summary(iv_qd, diagnostics = TRUE)
```

Inverse supply equation: p = f(qty, collusion, entry, month). To overcome endogeneity of qty, IV candidate is lakes. 
```{r}
iv.ps.po <- ivreg(log(gr) ~ log(tqg) + po + factor(era) + factor(month) 
              | lakes + po + factor(era) + factor(month), data = df)
iv.ps.pn <- ivreg(log(gr) ~ log(tqg) + pn + factor(era) + factor(month) 
              | lakes + pn + factor(era) + factor(month), data = df)

stargazer(qd, ps.po, ps.pn, iv.qd.po, iv.qd.pn, iv.ps.po, iv.ps.pn, type = "text")

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
