---
title: "Workshop 2"
author: "Filip Mellgren, David Hakula, Christopher Hayes"
date: '2019-04-10'
output:
  html_document:
    code_folding: hide
    df_print: kable
    highlight: zenburn
    theme: readable
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)
```

## R Markdown

```{r libraries, include = FALSE }
library(tidyverse)
library(rio)
library(ggthemes)
library(MASS)
library(stats4)

```

```{r import}
df <- import("Hypermarkets.dta")
df <- df %>% as_tibble
```

# Premises of the analysis 
Pros and cons of using only bix box stores
Argue about what the market is. Go to: lecture 2.
* SSNIP
* Other tests

Google scholar for sources

# Distribution of bib box stores



```{r}
df %>% group_by(stores) %>% 
  summarise(n = n()) %>% mutate(freq = n/ sum(n), cum = cumsum(n)/sum(n)) %>%
  t()
```



# Graphically illustrate 

```{r}
df %>% ggplot(aes(x = stores, y = pop)) +
  geom_point(alpha=0.5,  position = "jitter",shape = 19) +
  geom_point(alpha=0.5,  color = "red",shape = 19) +
  theme_economist() + scale_fill_economist() + scale_color_economist() +
  labs(title = "Heteroskedastic relationship", 
       caption = "A small amount of random noise was added to each point to better convey intuition about the conditional distribution",
  tag = "Graph 1", x = "Number of stores", y = "Population (1000's)", 
  color = "Stores", fill = "Stores")
```
Show a histogram of local market population across markets with different number of hy-
permarkets.
```{r density}
df %>% ggplot(aes(x = pop, fill = as.character(stores),
                  color = as.character(stores) )) +
  geom_histogram(alpha=0.5) + scale_fill_economist() + scale_color_economist() +
  theme_economist() +
  labs(title = "Large municipalities have more stores", 
       subtitle = "Yet there exists small towns with many stores", 
       caption = "A caption can be added",
  tag = "Graph 1", x = "Population (1000's)", y = "Density", 
  color = "Stores", fill = "Stores")
```
```{r save_plots}
word_margin <- 6
ggsave("Store_density_plot.png", height = 10, width = 29.7 - word_margin, units = "cm")
```

# 4 
```{r}
df <- as_tibble(df)
# pos_gpop and neg_gpop not included
# Include sd
tmp <- sapply(dplyr::select(df, stores, pop,
                     s_kids, s_pens, s_young, s_women, consumption,
                     hprice, hsold, wage, dist), summary)%>% t()
 
# Add correlation with stores as a sep. column

```



# 5 Variables included in the analysis
Y: pop, dist, 
X: gpop, pos_gpop, neg_gpop, s_kids, s_pens, s_young, s_women, consumption
W: hprice, hsold, wage, 


* What to include in S, V and F? Motivate why we include these variables
* Discuss but iterate a bit 

#### 6
Do file with the estimateion.

Ordered probit, what's the probability to observe 1, 2, 3 ,etc...
Important to specify probabilities to plug into likelihood function and plugginf in parameters that maximise probaility of observing the data.

Store in market if profit larger than 0. Observed variable is whether town in store, i.e. profit > 0. The latent variable is the profit which we don't observe.

Values she got:
pop: 1.36 and hprice: -.267 cons: .42

Note, these might have been an ordinary probit, as she spoke about the ordered probit only later.

Ordered probit
"Standard notes" from any text book in econometrics
```{r}
library(MASS)
polr(as.factor(stores) ~ pop + hprice, data=df, method = "probit")
 # results not exactly the same. Missing something?
```
Bresnahan and Reiss version. Slide 22 lecture 3 entry and market structure.
Profit decrease in n. Latent varibale is $\pi_n = - \beta x - \alpha_n$. y is number of stores.
$Pr(N = 0 \vert x) = Pr(\beta x - \alpha_1 + \varepsilon < 0) = 1 - \Phi(\bar{\pi_1})$
$Pr(N = 1 \vert x) = Pr(\beta x - \alpha_1 + \varepsilon > 0; \beta x - \alpha_2 + \varepsilon < 0) = \Phi(\bar{\pi}_1) - \Phi(\bar{\pi}_2)$.
```{r}
library(MASS)
polr(as.factor(stores) ~  pop + hprice, data=df, method = "probit")
```

gama fixed cost. Later entrants have higher fixed cost. 

arguments: a1 to a5, g1 to g5, S V and F

Only a1 and g1 to calculate profit1. Then we add more and more for profit_,2,3,4,5.

$V = \alpha_1 + X \beta - \sum_{n = 2}^N \alpha_N: a_1 - a_5$
$F = \gamma_1 + \gamma_L W_L + \sum_{n = 2}^N \gamma_N: g1 - g5$

Program gives all a_i, g_i, variable profit shifters betas: (young, ) cost shifters lambdas: (-1.84, 3.96), 
# WIll get results similar to what Bresnahan and Reiss have.

offset(pop) sets coef to 1.

Idea: Add a_1, a_2, ... a_5, g_1, ... g_5 as dummies for each store value


## Post lecture code
$\pi_{Nm} = S(Y, \lambda) V_{Nm}(X, \alpha, \beta) - F_{Nm}(W, \gamma) + \epsilon_m = \tilde{\pi}_{Nm} + \epsilon_m$

$V_{Nm} = \alpha_1 + \textbf{X}\beta  - \sum_{n = 2}^{N_m} \alpha_n$

$F_{Nm} = \gamma_1 + \gamma_L W_L +\sum_{n = 2}^{N_m} \gamma_n$

```{r}
# "Market size is assumed to be a linear function of population variables, where the 
#coefficient on population equals 1"

# S p.990
# town population + nearby population
#+ positive growth + negative growth
#+  commuters out of the county. 
# Twon population coefficient set to 1!!

# Generate all variables that we need:
df <- df %>% mutate(
              a1 = 1,
              a2 = if_else(stores >= 2, 1, 0),
              a3 = if_else(stores >= 3, 1, 0),
              a4 = if_else(stores >= 4, 1, 0),
              a5 = if_else(stores >= 5, 1, 0),
              g1 = 1,
              g2 = if_else(stores >= 2, 1, 0),
              g3 = if_else(stores >= 3, 1, 0),
              g4 = if_else(stores >= 4, 1, 0),
              g5 = if_else(stores >= 5, 1, 0))

# W is cost shifters
# hprice, p992 B & reiss

# X is local market characteristics
#per capita income in each industry's
#specification because consumer income usually affects the demand
#for goods and services. We included the number of births and the
#number of elderly residents in both doctors' and dentists' profit functions to control for #demographic variation in the demand for and
#cost of health care services.11 Because these variables summarize both
#demand and cost conditions, we do not attempt to draw structural
#inferences about the signs of their coefficients


# Y contain population variables
```

```{r}

df2 <- df %>% filter(stores > 0)
formula <- as.formula("as.ordered(stores) ~ (offset(pop) + pos_gpop + neg_gpop):(a1 + wage + a2 + a3 + a4 + a5) + hprice + g2 + g3 + g4 + g5")

formula <- as.formula("as.ordered(stores) ~ (offset(pop) + pos_gpop + neg_gpop):(wage + a1 + a2 )")

+ a3 + a4 + a5) + hprice + g2 + g3 + g4 + g5

polr(formula, data = df, lower = c(-Inf, -Inf, -Inf, 0.1, 0.1), upper = c(Inf, Inf, Inf, 1.01,1))
  
tmp3 <- polr(formula, data=df,method = "probit", na.action = na.omit)
```
```{r}

profit = function(a1, a2, a3, a4, a5, g1, g2, g3, g4, g5, gL){
  S <- -0.2
  V = 0.6
df <- df %>% mutate(
  S = pop * lambda_1, 
  P1 = pnorm( S * (V + a1) - g1 - gL*hprice), 
  P2 = pnorm(S * (V + a1 - a2) - gL*hprice - g1 - g2), 
  P3 = pnorm(S * (V + a1 - a2 - a3) - gL*hprice - g1 - g2 - g3), 
  P4 = pnorm(S * (V + a1 - a2 - a3 - a4) - gL*hprice - g1 - g2 - g3 - g4),
  P5 = pnorm(S * (V + a1 - a2 - a3 - a4 - a5) - gL*hprice - g1 - g2 - g3 - g4 - g5)
              )
df <- df %>% mutate(lnf = case_when(
  stores == 0 ~ log(1 - P1),
  stores == 1 ~ log(P1 -P2),
  stores == 2 ~ log(P2 -P3),
  stores == 3 ~ log(P3 -P4),
  stores == 4 ~ log(P4 -P5),
  stores == 5 ~log(P5),
  TRUE ~ 0))
  
# return the negative to maximize rather than minimize
-sum(df$lnf)
}

mle(profit, start = list(a1 = 0.7, a2 = 0.1, a3 = 0.1, a4 = 0.1, a5 = 0.1,
                         g1 = 0.1, g2 = 0.1, g3 = 0.1, g4 = 0.1, g5 = 0.1, gL = 0.1), 
    method = "L-BFGS-B", lower = c(0,0,0,0,0,0,0,0,0,0,-1),
      upper = c(1,1,1,1,1,1,1,1,1,1,1))

# TODO: set better interval for gL
# Include more in S, V
# 

```



